Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-05-29 overview.tex
removing an obsolete note
Adding my edits to the overview section.

\begin{eqnarray} 
\frac{\partial \rho}{\partial t} 
          + \frac{1}{a} \nabla \cdot (\rho \vecv) & = & 0
        \label{eq:mass} \\
%
\frac{\partial \rho \vecv}{\partial t}  
   + \frac{1}{a} \nabla \cdot \left(\rho \vecv \vecv + \myvec{I}p^* - \frac{\vecB \vecB}{a} \right) & = &
    - \frac{\dot{a}}{a} \rho \vecv - \frac{1}{a} \rho \nabla \phi
    %
%          + \frac{1}{a} ( \vecv \cdot \nabla ) \vecv
%    & = & - \frac{\dot{a}}{a} \vecv 
%          - \frac{1}{a \rho} \nabla p
%          - \frac{1}{a \rho} (\nalba \times \vecB) \times \vecB
%          - \frac{1}{a} \nabla \phi, 
        \label{eq:momentum} \\
%
\frac{\partial E}{\partial t} 
+ \frac{1}{a} \nabla \cdot \left[ (E + p^*) \vecv -  \frac{1}{a} \vecB (\vecB \cdot \vecv) \right] & = &
     - \frac{\dot{a}}{a} \left( 2E - \frac{B^2}{2a} \right)  - \frac{1}{a} \vecv \cdot \nabla \phi 
     - \Lambda + \Gamma + \frac{1}{a^2} \nabla \cdot \fcond
     %
%         + \frac{1}{a} \vecv \cdot \nabla E
 %  & = & - \frac{\dot{a}}{a} (3 \frac{p}{\rho} + {\vecv \cdot \vecv} )
  %       - \frac{1}{a \rho } \nabla \cdot (p \vecv) 
%%\nonumber   & &   %uncomment for 2-column
 %        - \frac{1}{a} \vecv \cdot \nabla \phi 
 %               + \Gamma - \Lambda.
       \label{eq:total_energy}  \\
%
\frac{\partial \vecB}{\partial t} - \frac{1}{a}  \nabla \times (\vecv \times \vecB) & = & 0 \label{eq:induction}
\end{eqnarray}

@@

The total fluid energy $E$ is given by:
\begin{equation}
E =  e + \frac{\rho v^2}{2}  + \frac{B^2}{2a}
        \label{eq:total_energy_def}.
\end{equation}
where $e$ is the comoving thermal energy density, and the total isotropic pressure $p^*$ is given by
\begin{equation}
p^* = p + \frac{B^2}{2a}
\end{equation}
The equations are closed with the equation of state and Poisson's equation (we work in the Newtonian limit):
%
\begin{eqnarray}
%
e   & = & p / [(\gamma - 1) \rho ],
        \label{eq:eq_of_state} \\
%
\nabla^2 \phi & = & 4 \pi G \rho_{\rm total}
%                   + 3 p_{total}/c^2) - \Lambda. 
        \label{eq:potential}
\end{eqnarray}
%
The equation of state is shown here for an ideal gas with a ratio of specific heats $\gamma$.
Equation (\ref{eq:potential}) includes contributions to the
gravitational potential $\phi$ from the total mass density $\rho_{\rm
  total}$, which is defined as $\rho_{\rm gas} + \rho_{\rm dm} + \rho_{\rm stars}
- \rho_{\rm average}$.  Although we write the equations including both MHD and
comoving coordinates, the code is frequently used both in the pure
hydrodynamic limit ($\vecB = 0$) and without cosmological expansion
($a = 1$, $\dot{a} = 0$).

@@

\begin{eqnarray}
\vecx & = & \vecr/a, \\
\vecv & = & a \hspace{0.5mm} d\vecx/dt = 
              \vecvp - \dot{a}\vecx, \\
\rho    & = & a^3 \rhop,   \\
p       & = & a^3 \pp, \\
E       & = & a^3 (\Ep - 
              \dot{a} \vecx \cdot \vecvp - 
              \frac{1}{2} \dot{a}^2 \vecx^2), \\
\phi    & = & \phip + \frac{1}{2} a \ddot{a} \vecx^2. \\
\vecB & = & a^2 \vecB^\prime
\end{eqnarray}
where primes indicate the usual quantities defined in a fixed frame frame, $\vecr$ is the proper distance and $\vecx$ is the comoving distance.

@@

The evolution of $a(t)$ is governed by the formula for the expansion of an isotropic, homogeneous universe:
%
\begin{eqnarray}
\frac{\ddot{a}}{a} & = & 
      - \frac{4 \pi G }{3 a^3 } (\rho_0 
      + 3p_0/c^2) 
      + \Lambda_c /3 .
      \label{eq:expansion} 
\end{eqnarray}
%
Here, $\rho_0$ is the mean comoving mass density (including both baryonic and dark matter), $p_0$ is the comoving background pressure contribution, and $\Lambda_c$ is the cosmological constant.
%$\gamma$ is the ratio of specific heats, and $\Lambda$ is the cosmological constant.
%We note that if $\gamma$ is not $5/3$ then $E$ in the first term on
%the right side of eq.~(ref{eq:total_energy}) must be replaced with 
%eq.~(\ref{eq:state}).
This system of equations is limited to the non-relativistic regime and assumes that curvature effects are not important --- both assumptions are reasonable as long as the  size of the simulated region is small compared to the radius of curvature and the Hubble length $c/H$ ($c$ is the speed of light and $H$ is the Hubble constant).

@@

Any collisionless component (e.g. dark matter, stars) is modeled by particles which are governed by Newton's equations in comoving coordinates:
%
\begin{eqnarray}
\frac{d \vecx}{dt} 
    & = & \frac{1}{a} \vecv, 
          \label{eq:dm_position} \\
\frac{d \vecv}{dt} 
    & = & - \frac{\dot{a}}{a} \vecv
          - \frac{1}{a} \nabla \phi, 
          \label{eq:dm_velocity} 
\end{eqnarray}
%
The particles also make a contribution to the gravitational potential through their contribution to Poisson's equation.

@@

In addition, we can solve the mass conservation equations for a set of chemical species and reactions.  For any species $i$ with comoving number density $n_i$, these equations have the form:
\begin{equation}
\frac{\partial n_i}{\partial t} 
          + \frac{1}{a} \nabla \cdot (n_i \vecv) = 
        k_{ij}(T) n_i n_j 
      + \Gamma^{ph}_j n_j 
        \label{eq:species_evolution}
\end{equation}
where $k_{ij}$ are the rate coefficients for the two-body reactions
and are usually functions of only the gas species (we will
specifically note the cases where we either include three-body
reactions or have density-dependent rates).  The $\Gamma^{ph}_j$ are
destruction/creation rates due to photoionizations and/or
photodissociations.  Currently the species we can follow include H,
H$^+$, He, He$^+$, He$^{++}$, and $e^-$, as well as additional options
that add H$^-$, H$_2$, H$_2^+$, and  HD, D, and D$^+$.

@@

The radiative transfer equation in comoving coordinates 
\citep[e.g.,][]{Gnedin97} reads
%
\begin{equation}
  \label{eq:rteqn}
  \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
  \frac{\hat{n} \cdot \nabla I_\nu}{\bar{a}} -
  \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
  3 I_\nu \right) = -\kappa_\nu I_\nu + j_\nu .
\end{equation}
%
Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is the radiation
specific intensity in units of energy per time $t$ per solid angle per
unit area per frequency $\nu$.  $H = \dot{a}/a$ is the Hubble
constant, where $a$ is the scale factor.  $\bar{a} = a/a_{em}$ is the
ratio of scale factors at the current time and time of emission.  The
second term represents the propagation of radiation, where the factor
$1/\bar{a}$ accounts for cosmic expansion.  The third term describes
both the cosmological redshift and dilution of radiation.  On the
right hand side, the first term considers the absorption coefficient
$\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$, and the second term
$j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ is the emission coefficient
that includes any sources of radiation (whether point or diffuse).

@@

As an alternative to the ray-casting strategy, we solve an equation derived from integrating
the equation of radiative transfer over angles.  In particular, we use a
flux-limited diffusion approximation for radiative
transfer, with couplings to both the gas energy and chemical number
densities. The equations solved are:
\begin{eqnarray}
  \label{eq:fld_radiation}
  \frac{\partial E_r}{\partial t} + \frac1a \nabla\cdot\left(E_r \vecv \right) &=& 
    \nabla\cdot\left(D\nabla E_r\right) -
    \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
  \label{eq:fld_heating}
  \frac{\partial e_c}{\partial t} &=& -\frac{2\dot{a}}{a} e_c + \Gamma - \Lambda,
%  \label{eq:fld_chemistry}
%  \partial_t {\mathrm n}_i + 
%    \frac1a \nabla\cdot\left({\mathrm n}_i \vecv_b\right) &=& 
%    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
%    \Gamma_i^{ph}, \quad i=1,\ldots,N_s,
\end{eqnarray}
where $E_r = E_r(x,t)$ is a grey radiation energy density and $e_c$ is the
internal energy correction due to photo-heating and chemical cooling.
%and ${\mathrm n}_i$ correspond to the chemical number densities of
%HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
%number density.  
Here, we define the grey radiation energy density
through first assuming a fixed frequency spectrum, i.e.
$E_{\nu}(\nu,\vecx,t) = \tilde{E}_r(\vecx,t) \chi(\nu)$, and then
defining the integrated quantity
\begin{equation}
\label{eq:grey_radiation_energy}
   E_r(\vecx,t) \equiv \int_{\nu_0}^{\infty}
   E_{\nu}(\nu,\vecx,t)\,\mathrm d\nu \  = \ 
   \tilde{E}_r(\vecx,t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
\end{equation}
In addition, $D$ in (\ref{eq:fld_radiation}) is the {\em Larsen
square-root flux-limiter} \citep[see][]{Morel2000}, $\kappa$ is the total
opacity, $\eta$ is the field of radiation sources, $\Gamma$ provides the
radiation induced photo-heating, and $\Lambda$ the chemical
cooling.
%, $\alpha_{i,j}$ are the reaction rate coefficients defining
%the interactions between species ${\mathrm n}_i$ and ${\mathrm n}_j$,
%and $\Gamma_i^{ph}$ are the radiation induced photo-ionization rates.

@@

Enzo implements the equations of isotropic heat conduction in a manner
similar to that of \citet{2007ApJ...664..135P}, where we model the
flux of isotropic heat transport:

\begin{equation}
\fcond = -\kappa_{sp} \nabla T
\end{equation}

Where $\kappa_{sp}$ is the Spitzer conduction coefficient
\citep{1962pfig.book.....S} and T is the baryon temperature (with
fluids explicitly assumed to be single-temperature).  Saturation of the heat flux in
high-temperature, low-density regimes (such as the intracluster medium
in galaxy clusters) is taken into account.  Thermal conduction in a
plasma can be strongly affected by the presence of magnetic field
lines, which may strongly suppress heat flow perpendicular to the
magnetic field.  In that case, we allow for heat transport only along
the magnetic field lines in simulations that include this physics, as follows:

@@

Given the hierarchy at some time $t$, we advance the solution in the
manner of a W-cycle in a multigrid solver.  First, we determine the
maximum time step allowed for the coarsest grid based on a variety of
accuracy and stability criteria and advance the grid by that time
interval, $\Delta t_0$.  We then move down to the next level and
advance all the grids on that level by a timestep $\Delta t_1$
($\Delta t_1 \leq \Delta t_0$), which is the minimum of all the allowed
timesteps for those grids.  If there are more levels, we repeat this
procedure until the bottom level of the hierarchy has been reached.
Once there, we continue advancing the grids on the lowest level until
they have `caught up' to the coarser parent grids (i.e. $\sum
\Delta t_l = \Delta t_{l-1}$).  This procedure repeats itself until
all grids have been advanced by a total time of $\Delta t_0$ (see
Figure~\ref{fig:wcycle} for a schematic).

Since interesting regions on the grid may move, the hierarchy must
adapt itself.  We do this whenever a level has caught up to the
coarser parent level and consists of entirely rebuilding the grids
on that level and at finer resolutions.  This is done by applying the grid refinement
criteria to the grids on that level, flagging zones that require
extra grids.  These criteria depend on the physical problem being
simulated -- see Section~\ref{sec:refinement_criteria} for more details.  Once a grid has a set of flagged cells,
we run a machine-vision based algorithm \citep{Berger91} to find edges
and determine a good placement of subgrids. 
%  These subgrids must not overlap one another, must cover all flagged
% cells and their neighbouring cells, and be above a preset efficiency
% threshold, where the efficiency of a cell is defined as the ratio of
% flagged cells to total cells. 
Once these new subgrids have been identified, the
solution from the next coarser grid is interpolated (see below) in
order to initialize the values on the new grids.  Finally, any overlap
between these new subgrids and the old ones is identified and the
solution within the regions of overlap is copied to the new subgrids.
The entire procedure just outlined is then repeated on the new grids
and in this way the entire hierarchy (from the original level examined
and continuing on to finer levels) is rebuilt.

@@

Four different (magneto)-hydrodynamic methods are implemented in
\enzo: (i) the hydrodynamic-only piecewise parabolic method (PPM)
developed by~\citet{1984JCoPh..54..174C} and extended to cosmology
by~\citet{1995CoPhC..89..149B}; (ii) the MUSCL-like Gudonov scheme
\citep{MUSCL} with or without magnetic fields and Dedner-based
divergence cleaning, described in more detail in
\citet{WangAbelZhang08} and \citet{WangAbel09}; (iii) a constrained
transport staggered MHD scheme \citep{Collins10}, and (iv) the
second-order finite-difference hydrodynamics method described in
\zeus~\citep{Stone92a,Stone92b}.

@@

This PPM implementation does an excellent job of capturing strong
shocks in a few cells.  Multidimensional schemes are built up by
directional splitting and produce a method that is formally second
order-accurate in space and time and which explicitly conserves mass,
linear momentum, and energy.  \red{(REF?)}  A variety of reconstruction methods and
Riemann solvers have been implemented.  \red{(These are never described!)}

@@

\subsubsection{Godunov MUSCL (HD) with Dedner divergence cleaning (MHD)}
This solver was developed by Wang \& Abel to attack problems in
magnetic field amplification during the formation of galaxies
\citep{Wang:2009a} and to understand the role of proto-stellar jets for
the theory of star formation \citep{Wang:2009b}. It combines the standard
approach of Godunov \citep{Godunov1959} for finite volume techniques
with the method of lines as described by
\cite{leveque2002finite} and \cite{toro-1997}. In addition, 
it implements the hyperbolic divergence cleaning algorithm of
\cite{2002JCoPh.175..645D}. It supports multiple approximate Riemann
solvers and non-ideal equations of states. Consequently,
these suite of solvers can be used for hydro and magneto-hydrodynamic
simulations. It is also this class of solvers as well as a version of
the PPM hydro solver which have been ported to Nvidia's CUDA framework
allowing \enzo\ to take advantage of modern graphics hardware \citep{Wang:2010}.

@@

The ZEUS method uses a staggered mesh such that the velocities are
face-centered, while the density and internal energy quantities are
cell-centered.  It splits the solution up into two steps, the
so-called source step, in which the momentum and energy values are
updated to reflect the pressure and gravity forces, as well as the
effect of an artificial viscosity required for stability.  The second
step (known as the transport step) accounts for the advection of
conserved quantities (mass, momentum and energy) across the grid.

@@

Collisionless matter (dark matter, stars, etc.) is modeled with
particles that interact with the baryons only via gravity.  These
particles are advanced in a single time step using a drift-kick-drift
algorithm\citep{Hockney88} to provide second-order accuracy even in the presence of
changing timesteps.  Since the particles follow the collapse of
structure, they are not adaptively refined.  Nor are there duplicate
sets of particles for each level; instead, each particle is associated
with the highest refined level available at its position in space, and is
moved as the hierarchy is rebuilt.  Thus, a particle has the same
timestep and feels the same gravitational force as a grid at that refinement
level.

@@

\enzo\ includes the capability of following up to
12 chemical species using a non-equilibrium solver.  The species can
be turned on in sets, with the simplest model including just H, H$^+$, 
He, He$^+$, He$^++$, and $e^-$, and more complete models adding first
species important for gas-phase molecular hydrogen formation: $H^-$,
H$_2$ and H$_2^+$, and then $HD$ formation: $HD$, $D$, $D+$.  (We note
that a reduced version of the chemistry, used only with the
flux-limited diffusion radiation transport, includes only hydrogen
species.)  The
cooling and heating due to these species is also included (see the
next section).  The solution of the rate equations is carried out
using one Jacobi iteration of an implicit Euler time discretization to
ensure stability; to
ensure accuracy we sub-cycle the rate equations with a shorter
timestep such that the electron and neutral fractions do not change by
more than 10\%.

@@

The code includes a photon-conserving radiative transfer algorithm
that is based on an adaptive ray-tracing method that utilizes the
HEALPix pixilation of the sphere \citep{AbelWandelt02}.  Photons are
integrated outwards from sources using an adaptive time-stepping
scheme that preserves accuracy in ionization fronts even in the
optically-thin limit.  This has been integrated with the chemistry and
cooling network to provide ionization and heating rates on a
cell-by-cell basis.  The method is described in more detail in our
main methodology section but is fully detailed and tested in
\citet{Wise11_Moray}.

>>>>

\begin{eqnarray} 
  \frac{\partial \rho}{\partial t} 
  + \frac{1}{a} \nabla \cdot (\rho \vecv) & = & 0
  \label{eq:mass} \\
%
  \frac{\partial \rho \vecv}{\partial t}  
  + \frac{1}{a} \nabla \cdot \left(\rho \vecv \vecv + \myvec{I}p^* - 
    \frac{\vecB \vecB}{a} \right) & = &
  - \frac{\dot{a}}{a} \rho \vecv - \frac{1}{a} \rho \nabla \phi
% 
%          + \frac{1}{a} ( \vecv \cdot \nabla ) \vecv
%    & = & - \frac{\dot{a}}{a} \vecv 
%          - \frac{1}{a \rho} \nabla p
%          - \frac{1}{a \rho} (\nalba \times \vecB) \times \vecB
%          - \frac{1}{a} \nabla \phi, 
  \label{eq:momentum} \\
%
  \frac{\partial E}{\partial t} 
  + \frac{1}{a} \nabla \cdot \left[ (E + p^*) \vecv - 
    \frac{1}{a} \vecB (\vecB \cdot \vecv) \right] & = &
  - \frac{\dot{a}}{a} \left( 2E - \frac{B^2}{2a} \right) - 
  \frac{1}{a} \vecv \cdot \nabla \phi 
  - \Lambda + \Gamma + \frac{1}{a^2} \nabla \cdot \fcond
% 
%         + \frac{1}{a} \vecv \cdot \nabla E
%  & = & - \frac{\dot{a}}{a} (3 \frac{p}{\rho} + {\vecv \cdot \vecv} )
%       - \frac{1}{a \rho } \nabla \cdot (p \vecv) 
%%\nonumber   & &   %uncomment for 2-column
%        - \frac{1}{a} \vecv \cdot \nabla \phi 
%               + \Gamma - \Lambda.
  \label{eq:total_energy}  \\
%
  \frac{\partial \vecB}{\partial t} - 
  \frac{1}{a}  \nabla \times (\vecv \times \vecB) & = & 
  -\frac{\dot{a}}{2 a} \vecB \label{eq:induction}
\end{eqnarray}

@@

The total fluid energy $E$ is given by:
\begin{equation}
E =  e + \frac{\rho v^2}{2}  + \frac{B^2}{2a}
        \label{eq:total_energy_def}.
\end{equation}
where $e$ is the comoving thermal energy density, and the total
isotropic pressure $p^*$ is given by
\begin{equation}
p^* = p + \frac{B^2}{2a},
\end{equation}
where $p$ is the gas pressure.  The equations are closed with the
equation of state and Poisson's equation (we work in the Newtonian
limit):
%
\begin{eqnarray}
%
  e &=& p / [(\gamma - 1) \rho ],
  \label{eq:eq_of_state} \\
%
  \nabla^2 \phi &=& 4 \pi G \rho_{\rm total}
%                   + 3 p_{total}/c^2) - \Lambda. 
  \label{eq:potential}
\end{eqnarray}
%
The equation of state is shown here for an ideal gas with a ratio of
specific heats $\gamma$.  Equation (\ref{eq:potential}) includes
contributions to the gravitational potential $\phi$ from the total
mass density $\rho_{\rm total}$, which is defined as $\rho_{\rm gas} +
\rho_{\rm dm} + \rho_{\rm stars} - \rho_{\rm average}$.  Although we
write the equations including both MHD and comoving coordinates, the
code is frequently used both in the pure hydrodynamic limit ($\vecB =
0$) and without cosmological expansion ($a = 1$, $\dot{a} = 0$).

@@

\begin{eqnarray}
\vecx & = & \vecr/a, \\
\vecv & = & a \hspace{0.5mm} d\vecx/dt = 
              \vecvp - \dot{a}\vecx, \\
\rho  & = & a^3 \rhop,   \\
p     & = & a^3 \pp, \\
E     & = & a^3 \left(\Ep - 
            \dot{a} \vecx \cdot \vecvp - 
            \frac{1}{2} \dot{a}^2 \vecx^2\right), \\
\phi  & = & \phip + \frac{1}{2} a \ddot{a} \vecx^2. \\
\vecB & = & a^2 \vecB^\prime,
\end{eqnarray}
where primes indicate the usual quantities defined in a fixed frame
frame, $\vecr$ is the proper distance and $\vecx$ is the comoving
distance.

@@

The evolution of $a(t)$ is governed by the formula for the expansion
of an isotropic, homogeneous universe:
%
\begin{eqnarray}
\frac{\ddot{a}}{a} & = & 
      - \frac{4 \pi G }{3 a^3 } (\rho_0 
      + 3p_0/c^2) 
      + \Lambda_c /3 .
      \label{eq:expansion} 
\end{eqnarray}
%
Here, $\rho_0$ is the mean comoving mass density (including both
baryonic and dark matter), $p_0$ is the comoving background pressure
contribution, and $\Lambda_c$ is the cosmological constant.
%$\gamma$ is the ratio of specific heats, and $\Lambda$ is the cosmological constant.
%We note that if $\gamma$ is not $5/3$ then $E$ in the first term on
%the right side of eq.~(ref{eq:total_energy}) must be replaced with 
%eq.~(\ref{eq:state}).
This system of equations is limited to the non-relativistic regime and
assumes that curvature effects are not important --- both assumptions
are reasonable as long as the size of the simulated region is small
compared to the radius of curvature and the Hubble length $c/H$, where
$c$ is the speed of light and $H = \dot{a}/a$ is the Hubble constant.

@@

Any collisionless component (e.g., dark matter, stars) is modeled by
particles which are governed by Newton's equations in comoving
coordinates:
%
\begin{eqnarray}
\frac{d \vecx}{dt} 
    & = & \frac{1}{a} \vecv, 
          \label{eq:dm_position} \\
\frac{d \vecv}{dt} 
    & = & - \frac{\dot{a}}{a} \vecv
          - \frac{1}{a} \nabla \phi, 
          \label{eq:dm_velocity} 
\end{eqnarray}
%
The particles also make a contribution to the gravitational potential
through their contribution to Poisson's equation (Equation
\ref{eq:potential}).

@@

In addition, we can solve the mass conservation equations for a set of
chemical species and reactions.  For any species $i$ with comoving
number density $n_i$, these equations have the form:
\begin{equation}
  \frac{\partial n_i}{\partial t} 
  + \frac{1}{a} \nabla \cdot (n_i \vecv) = 
  k_{ij}(T) n_i n_j + \Gamma^{\rm ph}_j n_j 
  \label{eq:species_evolution}
\end{equation}
where $k_{ij}$ are the rate coefficients for the two-body reactions
and are usually functions of only the gas species (we will
specifically note the cases where we either include three-body
reactions or have density-dependent rates).  The $\Gamma^{\rm ph}_j$
are destruction/creation rates due to photoionizations and/or
photodissociations.  Currently the species we can follow include H,
H$^+$, He, He$^+$, He$^{++}$, and e$^-$, as well as additional options
that add H$^-$, H$_2$, H$_2^+$, and HD, D, and D$^+$.  We also can
track the advection of a comoving metal density that may contribute to
the radiative cooling.

@@

The radiative transfer equation in comoving coordinates
\citep[e.g.,][]{Gnedin97} reads
%
\begin{equation}
  \label{eq:rteqn}
  \frac{1}{c} \; \frac{\partial I_\nu}{\partial t} + 
  \frac{\hat{n} \cdot \nabla I_\nu}{\bar{a}} -
  \frac{H}{c} \; \left( \nu \frac{\partial I_\nu}{\partial \nu} -
  3 I_\nu \right) = -\kappa_\nu I_\nu + j_\nu .
\end{equation}
%
Here $I_\nu \equiv I(\nu, \mathbf{x}, \Omega, t)$ is the radiation
specific intensity in units of energy per time $t$ per solid angle per
unit area per frequency $\nu$.  $\bar{a} = a/a_{em}$ is the ratio of
scale factors at the current time and time of emission.  The second
term represents the propagation of radiation, where the factor
$1/\bar{a}$ accounts for cosmic expansion.  The third term describes
both the cosmological redshift and dilution of radiation.  On the
right hand side, the first term considers the absorption coefficient
$\kappa_\nu \equiv \kappa_\nu(\mathbf{x},\nu,t)$, and the second term
$j_\nu \equiv j_\nu(\mathbf{x},\nu,t)$ is the emission coefficient
that includes any sources of radiation (whether point or diffuse).

@@

As an alternative to the ray-casting strategy, we solve an equation
derived from integrating the equation of radiative transfer over
angles.  In particular, we use a flux-limited diffusion approximation
for radiative transfer, with couplings to both the gas energy and
chemical number densities. The equations solved are:
\begin{eqnarray}
  \label{eq:fld_radiation}
  \frac{\partial E_r}{\partial t} + \frac1a \nabla\cdot\left(E_r \vecv \right) &=& 
  \nabla\cdot\left(D\nabla E_r\right) -
  \frac{\dot{a}}{a} E_r - c\kappa E_r + \eta, \\
  \label{eq:fld_heating}
  \frac{\partial e_c}{\partial t} &=& -\frac{2\dot{a}}{a} e_c + \Gamma - \Lambda,
%  \label{eq:fld_chemistry}
%  \partial_t {\mathrm n}_i + 
%    \frac1a \nabla\cdot\left({\mathrm n}_i \vecv_b\right) &=& 
%    \alpha_{i,j} {\mathrm n}_e {\mathrm n}_j - {\mathrm n}_i
%    \Gamma_i^{ph}, \quad i=1,\ldots,N_s,
\end{eqnarray}
where $E_r = E_r(\vecx,t)$ is a grey radiation energy density and
$e_c$ is the internal energy correction due to photo-heating and
chemical cooling.
%and ${\mathrm n}_i$ correspond to the chemical number densities of
%HI, HII, HeI, HeII and HeIII, and ${\mathrm n}_e$ is the electron
%number density.  
Here, we define the grey radiation energy density through first
assuming a fixed frequency spectrum, i.e.  $E_{\nu}(\nu,\vecx,t) =
\tilde{E}_r(\vecx,t) \chi(\nu)$, and then defining the integrated
quantity
\begin{equation}
\label{eq:grey_radiation_energy}
   E_r(\vecx,t) \equiv \int_{\nu_0}^{\infty}
   E_{\nu}(\nu,\vecx,t)\,\mathrm d\nu \  = \ 
   \tilde{E}_r(\vecx,t) \int_{\nu_0}^{\infty} \chi(\nu)\,\mathrm d\nu.
\end{equation}
In addition, $D$ in Equation (\ref{eq:fld_radiation}) is the Larsen
square-root flux-limiter \citep[see][]{Morel2000}; $\kappa$ is the
total opacity; $\eta$ is the field of radiation sources; $\Gamma$
provides the radiation induced photo-heating, and $\Lambda$ the
chemical cooling.
%, $\alpha_{i,j}$ are the reaction rate coefficients defining
%the interactions between species ${\mathrm n}_i$ and ${\mathrm n}_j$,
%and $\Gamma_i^{ph}$ are the radiation induced photo-ionization rates.

@@

Enzo implements the equations of isotropic heat conduction in a manner
similar to that of \citet{2007ApJ...664..135P}, where we model the
flux of isotropic heat transport,
%
\begin{equation}
\fcond = -\kappa_{sp} \nabla T.
\end{equation}
%
Here $\kappa_{sp}$ is the Spitzer conduction coefficient
\citep{1962pfig.book.....S} and T is the baryon temperature (with
fluids explicitly assumed to be single-temperature).  Saturation of
the heat flux in high-temperature, low-density regimes (such as the
intracluster medium in galaxy clusters) is taken into account.
Thermal conduction in a plasma can be strongly affected by the
presence of magnetic field lines, which may strongly suppress heat
flow perpendicular to the magnetic field.  In that case, we allow for
heat transport only along the magnetic field lines in simulations that
include this physics, as follows:

@@

Given the hierarchy at some time $t$, we advance the solution in the
manner of a W-cycle in a multigrid solver.  First, we determine the
maximum time step allowed for the coarsest grid based on a variety of
accuracy and stability criteria and advance the grid by that time
interval, $\Delta t_0$.  We then move down to the next level and
advance all the grids on that level by a timestep $\Delta t_1$
($\Delta t_1 \leq \Delta t_0$), which is the minimum of all the
allowed timesteps for those grids.  If there are more levels, we
repeat this procedure until the bottom level of the hierarchy has been
reached.  Once there, we continue advancing the grids on the lowest
level until they have ``caught up'' to the coarser parent grids
(i.e. $\sum \Delta t_l = \Delta t_{l-1}$).  This procedure repeats
itself until all grids have been advanced by a total time of $\Delta
t_0$ (see Figure~\ref{fig:wcycle} for a schematic).

\begin{figure}
\begin{center}
%\includegraphics[width=0.6\textwidth]{figures/wcycle.eps}
\includegraphics[width=0.6\textwidth]{figures/timestepping.eps}
\end{center}
\caption{\emph{Left:} Example of the timesteps on a 2-level AMR
  hierarchy.  Enzo does not restrict the timesteps on the finer levels
  to be a factor of $1/2^n$ smaller.  \emph{Right:} The order in which
  the AMR grids are evaluated on each level.\vspace{1ex}}
\label{fig:wcycle}
\end{figure}

Since interesting regions on the grid may move, the hierarchy must
adapt itself.  We do this whenever a level has caught up to the
coarser parent level and consists of entirely rebuilding the grids on
that level and at finer resolutions.  This is done by applying the
grid refinement criteria to the grids on that level, flagging zones
that require extra grids.  These criteria depend on the physical
problem being simulated -- see Section~\ref{sec:refinement_criteria}
for more details.  Once a grid has a set of flagged cells, we run a
machine-vision based algorithm \citep{Berger91} to find edges and
determine a good placement of subgrids.
%  These subgrids must not overlap one another, must cover all flagged
% cells and their neighbouring cells, and be above a preset efficiency
% threshold, where the efficiency of a cell is defined as the ratio of
% flagged cells to total cells. 
Once these new subgrids have been identified, the solution from the
next coarser grid is interpolated in order to initialize the values on
the new grids.  Finally, any overlap between these new subgrids and
the old ones is identified, and the solution within the regions of
overlap is copied to the new subgrids.  The entire procedure just
outlined is then repeated on the new grids and in this way the entire
hierarchy (from the original level examined and continuing on to finer
levels) is rebuilt.

@@

Four different (magneto)-hydrodynamic methods are implemented in
\enzo: (i) the hydrodynamic-only piecewise parabolic method (PPM)
developed by~\citet{1984JCoPh..54..174C} and extended to cosmology
by~\citet{1995CoPhC..89..149B}; (ii) the MUSCL-like Godunov scheme
\citep{MUSCL} with or without magnetic fields and Dedner-based
divergence cleaning, described in more detail in
\citet{WangAbelZhang08} and \citet{WangAbel09}; (iii) a constrained
transport staggered MHD scheme \citep{Collins10}, and (iv) the
second-order finite-difference hydrodynamics method described in
\zeus~\citep{Stone92a,Stone92b}.

@@

This PPM implementation does an excellent job of capturing strong
shocks in a few cells.  Multidimensional schemes are built up by
directional splitting and produce a method that is formally second
order-accurate in space and time and which explicitly conserves mass,
linear momentum, and energy.  \red{(REF?)}  A variety of Riemann
solvers have been implemented.

@@

\subsubsection{Godunov MUSCL (HD) with Dedner divergence cleaning (MHD)}
This solver was developed to attack problems in magnetic field
amplification during the formation of galaxies \citep{Wang:2009a} and
to understand the role of proto-stellar jets for the theory of star
formation \citep{Wang:2009b}. It combines the standard approach of
Godunov \citep{Godunov1959} for finite volume techniques with the
method of lines as described by \cite{leveque2002finite} and
\cite{toro-1997}. In addition, it implements the hyperbolic divergence
cleaning algorithm of \cite{2002JCoPh.175..645D}. It supports multiple
approximate Riemann solvers and non-ideal equations of
states. Consequently, these suite of solvers can be used for hydro and
magneto-hydrodynamic simulations. It is also this class of solvers as
well as a version of the PPM hydro solver which have been ported to
Nvidia's CUDA framework allowing \enzo\ to take advantage of modern
graphics hardware \citep{Wang:2010}.

@@

The \zeus\ method uses a staggered mesh such that the velocities are
face-centered, while the density and internal energy quantities are
cell-centered.  It splits the solution up into two steps, the
so-called source step, in which the momentum and energy values are
updated to reflect the pressure and gravity forces, as well as the
effect of an artificial viscosity required for stability.  The second
step (known as the transport step) accounts for the advection of
conserved quantities (mass, momentum and energy) across the grid.

@@

Collisionless matter (e.g., dark matter, stars, etc.) is modeled with
particles that interact with the baryons only via gravity.  These
particles are advanced in a single time step using a drift-kick-drift
algorithm \citep{Hockney88} to provide second-order accuracy even in
the presence of changing timesteps.  Since the particles follow the
collapse of structure, they are not adaptively refined.  Nor are there
duplicate sets of particles for each level; instead, each particle is
associated with the highest refined level available at its position in
space and is moved as the hierarchy is rebuilt.  Thus, a particle has
the same timestep and feels the same gravitational force as a grid at
that refinement level.

@@

\enzo\ includes the capability of following up to 12 chemical species
using a non-equilibrium solver.  The species can be turned on in sets,
with the simplest model including just H, H$^+$, He, He$^+$, He$^{++}$,
and e$^-$, and more complete models adding first species important for
gas-phase molecular hydrogen formation: H$^-$, H$_2$ and H$_2^+$, and
then HD formation: HD, D, D$^+$.  (We note that a reduced version
of the chemistry, used only with the flux-limited diffusion radiation
transport, includes only hydrogen species.)  The cooling and heating
due to these species is also included (see the next section).  The
solution of the rate equations is carried out using one Jacobi
iteration of an implicit Euler time discretization to ensure
stability; to ensure accuracy we sub-cycle the rate equations with a
shorter timestep such that the electron and neutral fractions do not
change by more than 10\%.

@@

The code includes a photon-conserving radiative transfer algorithm
that is based on an adaptive ray-tracing method that utilizes the
HEALPix pixelization of a sphere \citep{AbelWandelt02}.  Photons are
integrated outwards from sources using an adaptive time-stepping
scheme that preserves accuracy in ionization fronts even in the
optically-thin limit.  This has been integrated with the chemistry and
cooling network to provide ionization and heating rates on a
cell-by-cell basis.  The method is described in more detail in our
main methodology section but is fully detailed and tested in
\citet{Wise11_Moray}.
